---
title: "Lab 4B: Foundations for statistical inference - Sampling distributions and Sample Size"
author: "Intro Stats"
date: "Fall 2018"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, results='hide', message=FALSE, warning=FALSE)
require(tidyverse)
ames <- read_csv("ames.csv")
```

## Sample size and the sampling distribution

The approximate sampling distribution that we computed tells us much about estimating 
the average living area in homes in Ames.  Because the sample mean is an 
*unbiased estimator*, the sampling distribution is centered at the true mean average 
living area of the the population, and the spread of the distribution 
indicates how much variability is induced by sampling only 50 home sales.

To get a sense of the effect that sample size has on our distribution, let's 
build up two more approximate sampling distributions: one based on a sample size of 10 and 
another based on a sample size of 100.

```{r samp-10-100, eval=FALSE}
sample_means10 <- 
  ames %>% 
  rep_sample_n(size = 10, reps = 5000) %>% 
  group_by(replicate) %>% 
  summarize(meanLivArea = mean(LivArea))

sample_means100 <- 
  ames %>% 
  rep_sample_n(size = 100, reps = 5000) %>% 
  group_by(replicate) %>% 
  summarize(meanLivArea = mean(LivArea))
```

To see the effect that different sample sizes have on the histograms of sample 
means, plot the three distributions one after the other.

```{r plot-samps, eval=FALSE, tidy = FALSE}
sample_means10 %>% 
  ggplot(aes(x = meanLivArea))+
  geom_histogram()

sample_means50 %>% 
  ggplot(aes(x = meanLivArea))+
  geom_histogram()

sample_means100 %>% 
  ggplot(aes(x = meanLivArea))+
  geom_histogram()
```

The `bins` argument specifies the number of bins used in constructing the histogram.

<!-- Cut the `xlim` argument bit - maybe they look at the tick marks instead to keep the code simpler?-->

6.  When the sample size is larger, what happens to the center?  the shape?  the spread?

* * *

### On your own

Open up a new R Script.

So far, we have only focused on estimating the mean living area in homes in 
Ames.  Now you'll try to estimate the mean home price.

-   Take a random sample of size 50 from `ames`. Using this sample, what is 
    your best point estimate of the population mean of `SalePrice`?

-   Since you have access to the population, simulate the approximate sampling 
    distribution for $\bar{x}_{price}$ by taking 5000 samples from the 
    population of size 50 and computing 5000 sample means.  Store these means 
    in a vector called `sample_means50`. Plot the data and  describe the 
    shape of this distribution. Based on this distribution of sample means, 
    what would you guess the mean home price of the population to be? Finally, 
    calculate and report the population mean.

-   Change your sample size from 50 to 150, then compute 5000 sample means using the same method as above, and store these means in a new vector called `sample_means150`. Describe the shape of this 
    distribution, and compare it to the distribution for a sample 
    size of 50.  Based on this distribution, what would you guess to 
    be the mean sale price of homes in Ames?

-   Of the approximate sampling distributions you created from sample sizes of 50 and 150, which has a smaller range?  If we're concerned with making estimates that are more often close to the true value, would we prefer a distribution with a large or small range?  Why?

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.  Modified by Ted Galanthay and Matt Thomas, July 2018.
</div>