---
title: "Lab 4B: Foundations for statistical inference - Sampling distributions and Sample Size"
author: "Intro Stats"
date: "Fall 2018"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, results='hide', message=FALSE, warning=FALSE)
require(tidyverse)
require(moderndive)
ames <- read_csv("ames.csv")
```

## Sample size and the sampling distribution

The approximate sampling distribution that we computed tells us much about estimating 
the average living area in homes in Ames.  Because the sample mean is an 
*unbiased estimator*, the sampling distribution is centered at the true mean average 
living area of the the population, and the spread of the distribution 
indicates how much variability is induced by sampling only 50 home sales.

To get a sense of the effect that sample size has on our distribution, let's 
build up two more approximate sampling distributions: one based on a sample size of 10 and 
another based on a sample size of 100. Make sure `tidyverse` and `moderndive` are loaded.

```{r samp-10-100, eval=FALSE}
sample_means10 <- 
  ames %>% 
  rep_sample_n(size = 10, reps = 5000) %>% 
  group_by(replicate) %>% 
  summarize(meanLivArea = mean(LivArea))

sample_means100 <- 
  ames %>% 
  rep_sample_n(size = 100, reps = 5000) %>% 
  group_by(replicate) %>% 
  summarize(meanLivArea = mean(LivArea))
```

Add another set of sample means, this time of size 50.

To see the effect that different sample sizes have on the histograms of sample 
means, plot the three distributions one after the other.

```{r plot-samps, eval=FALSE, tidy = FALSE}
sample_means10 %>% 
  ggplot(aes(x = meanLivArea))+
  geom_histogram()

sample_means50 %>% 
  ggplot(aes(x = meanLivArea))+
  geom_histogram()

sample_means100 %>% 
  ggplot(aes(x = meanLivArea))+
  geom_histogram()
```

(Note: The `bins` argument specifies the number of bins used in constructing the histogram.)

1. Include your 3 histograms.

2.  As the sample size gets larger, what happens to the center?  The shape?  The spread? (Note: pay attention to the x-axis values.)

3. Determine the standard deviation of the Living area for the population. How does it compare to the standard deviation of each of your sample_mean datasets. Try to come up with a formula relating them.

* * *

### On your own

Write your commands for this section in a new R Script.

- Now we'll keep the sample size the same (50) but change the number of samples. Create 3 data sets: one with 100 samples of size 50, one with 5000 samples of size 50, and one with 100000 samples of size 50. State the mean and standard deviation of each of them.

- How does the number of samples affect your sampling distribution? Comment on both the shape, center, and spread.

- Summarize the ideas of this lab. We have modified the sample size when constructing a sampling distribution as well as the number of samples. How do each of them relate to the sampling distribution?

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.  Modified by Ted Galanthay and Matt Thomas, August 2018.
</div>