---
title: 'Lab 5A: Categorical Data and Confidence Intervals'
author: "Intro Stats"
date: "Fall 2018"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, results='hide', message=FALSE, warning=FALSE)
require(tidyverse)
require(infer)
#require(moderndive)
atheism <- read_csv("atheism.csv")
```

In August of 2012, news outlets ranging from the [Washington
Post](http://www.washingtonpost.com/national/on-faith/poll-shows-atheism-on-the-rise-in-the-us/2012/08/13/90020fd6-e57d-11e1-9739-eef99c5fb285_story.html) to the [Huffington
Post](http://www.huffingtonpost.com/2012/08/14/atheism-rise-religiosity-decline-in-america_n_1777031.html) ran a story about the rise of atheism in America. The source for the story was 
a poll that asked people, "Irrespective of whether you attend a place of 
worship or not, would you say you are a religious person, not a religious 
person or a convinced atheist?" This type of question, which asks people to 
classify themselves in one way or another, is common in polling and generates 
categorical data. In this lab we take a look at the atheism survey and explore 
what's at play when making inference about population proportions using 
categorical data.

## The survey

To access the press release for the poll, conducted by WIN-Gallup 
International, click on the following link:

<https://web.archive.org/web/20121016062403/http://redcresearch.ie/wp-content/uploads/2012/08/RED-C-press-release-Religion-and-Atheism-25-7-12.pdf>

Take a moment to review the report then address the following questions.

1.  On the first page, several key findings are reported. Do these 
    percentages appear to be *sample statistics* (derived from the data 
    sample) or *population parameters*?

2.  The title of the report is "Global Index of Religiosity and Atheism". To
    generalize the report's findings to the global human population, what must 
    we assume about the sampling method? Does that seem like a reasonable 
    assumption?

## The data

Turn your attention to Table 6 (pages 14 and 15), which reports the
sample size and response percentages for all 57 countries. While this is
a useful format to summarize the data, we will base our analysis on the
original data set of individual responses to the survey. This data is in the `atheism.csv` file. Load this data set into R.

3.  What does each row of Table 6 correspond to? What does each row of 
    `atheism` correspond to?

To investigate the link between these two ways of organizing this data, take a 
look at the estimated proportion of atheists in the United States. Towards 
the bottom of Table 6, we see that this is 5%. We should be able to come to 
the same number using the `atheism` data.

4.  We can count the number of each response in each category:

```{r}
atheism %>% 
  group_by(nationality) %>% 
  count(response)
```

but this is quite difficult to read. We can first take a subset to only look at the data in the US in 2012. Using the command below, create a new dataframe called `us12` that contains only the rows in `atheism` associated with respondents to the 2012 survey from the United States. 

```{r eval=FALSE}
us12 <-
 atheism %>%
 filter(nationality == "United States",year == "2012")
```

We can also get a table of the responses:

```{r, eval=FALSE}
table(us12$response)
```

(An alternative would be to use a summarize, but this is quite quick.)

5. Calculate the proportion of atheist responses. Does it agree with the percentage in Table 6? If not, why?

## Inference on proportions

As was hinted at in Exercise 1, Table 6 provides *statistics*, that is, 
calculations made from the sample of 51,927 people. What we'd like, though, is 
insight into the population *parameters*. You answer the question, "What 
proportion of people in your sample reported being atheists?" with a 
statistic; while the question "What proportion of people on earth would report 
being atheists" is answered with an estimate of the parameter.

<!--
The inferential tools for estimating population proportion are analogous to 
those used for means: the confidence interval and the hypothesis test.
-->

6.  Write out the conditions for constructing a 95% confidence
    interval for the proportion of atheists in the United States in 2012.
    Are you confident all conditions are met?

If the conditions for inference are reasonable, we can either calculate the standard error and construct the interval by hand, or allow the `prop.test()` function to do it for us.

7. Construct a 95% confidence interval for the proportion of the US population in 2012 who were atheists.

8. Use the following R code to have R construct a confidence interval for you. You may have to hunt through the output to find the confidence interval.

```{r eval=FALSE}
prop.test(table(us12$response))

```


Note that since the goal is to construct an interval estimate for a 
proportion, it's necessary to specify what constitutes a "success", which here 
is a response of `"atheist"` as specified in the output.

Although formal confidence intervals and hypothesis tests don't show up in the 
report, suggestions of inference appear at the bottom of page 6: "In general, 
the error margin for surveys of this kind is $\pm$ 3-5% at 95% confidence".

7.  Based on the R output, what is the margin of error for the estimate of the 
    proportion of the proportion of atheists in US in 2012?

* * *

### On your own

* Using the `prop.test()` function, calculate confidence intervals for the 
    proportion of atheists in 2012 in two other countries of your choice, and 
    report the associated margins of error. Be sure to note whether the 
    conditions for inference are met. It may be helpful to create new data 
    sets for each of the two countries first, and then use these data sets in 
    the `prop.test()` function to construct the confidence intervals.


* Is there convincing evidence that Spain has seen a change in its 
    atheism index between 2005 and 2012?\
    *Hint:* Create a new data set for respondents from Spain. Form
    confidence intervals for the true proportion of athiests in both
    years, and determine whether they overlap.

* Is there convincing evidence that the United States has seen a
    change in its atheism index between 2005 and 2012?

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel. Modified by Ted Galanthay and Matt Thomas, November 2018.
</div>